Water desalination using polyelectrolyte hydrogel. Gibbs ensemble modelling

### 2022-06-27 gibbs.tex
LN: water desalination technologies

The first approach~--- distillation. It uses thermal means to effect
a phase change of the water to vapor (or solid), then to physically
separate the vapor phase from the remaining salt solution, and then
recover the thermal energy for reuse as the separated vapor reverts
to liquid form. Distillation processes were the first desalination
techniques conducted on a large commercial scale and still account
for a large portion of the modern world’s desalination capacity.

The second approach is to physically separate the brine components,
generally with a membrane, as they move in response to an externally
applied gradient. In the context of our study, we will mention the
reverse osmosis process (RO) \textasciitilde -{}-{}- the major processes
of all the modern desalination industry, and the new emerging membrane
technology -{}-{}- forward osmosis (FO) \cite{Akther2015}. In RO,
water passes through a membrane that is impermeable to the solute
in response to a chemical potential gradient achieved through pressurization.
In FO, a concentrated draw solution, with a lower water chemical potential,
and a feed solution, with a higher water chemical potential, are separated
by a membrane that rejects the salt but allows the passage of water.
The water permeation from feed solution to draw solution is a spontaneous
process driven by the chemical potential gradient.

Distillation is easy and cheap technology but it is characterized
by relatively high energy costs due to the dissipation of thermal
energy. On the contrary, RO uses expensive osmotic membranes, which
need to be replaced regularly because of scaling and fouling. Moreover,
RO requires very high operating pressures, ranging from 20 to 200
bar, to let the water pass through the membrane. But in terms of energy
losses, it works close to the thermodynamic limit. Thus, theoretically,
per one transferred from freshwater to salty water ion pair, it consumes
only the energy of difference of those ions chemical potentials.

The absence of large hydraulic pressure in FO process in comparison
to RO allows to reduce the energy consumption in pumping, reduce membrane
scaling and fouling, therefore significantly increase the lifetime
of the membranes. In FO, the draw solutes (agents) dispersed and/or
dissolved in water form homogeneous draw solutions. The correct choice
of the draw agents is a task of paramount importance. As an osmotically
driven process, the draw solute is expected to significantly reduce
the water chemical potential, and consequently generate a high osmotic
pressure. On the other hand, the diluted draw solution is also expected
to be easily separated to fresh water and to concentrated draw solution
for reuse \cite{Cai2016}.

>>>>

The first approach is distillation, which uses heat to cause
a phase change of the water to vapor. Then we physically
separate the vapor phase from the remaining salt solution, and after that we
recover the thermal energy for reuse as the separated vapor reverts back
to the liquid form. Distillation processes were the first desalination
techniques conducted on a large commercial scale and still account
for a large portion of the modern world’s desalination capacity.

The second approach is to physically separate the brine components
using a membrane, as they move in response to an externally
\todoi{we need to rephrase this}
applied gradient.
In the context of our study, we will mention the
reverse osmosis process (RO) \textasciitilde -{}-{}- the major processes
of all the modern desalination industry, and the newly emerging membrane
technology -{}-{}- forward osmosis (FO) \cite{Akther2015}. In RO,
water passes through a membrane that is impermeable to the solute
in response to a chemical potential gradient achieved through pressurization.
In FO, a concentrated draw solution, with a lower water chemical potential,
and a feed solution, with a higher water chemical potential, are separated
by a membrane that rejects the salt but allows the passage of water.
The water permeation from feed solution to draw solution is a spontaneous
process driven by the chemical potential gradient.

Distillation is easy and cheap technology but it is characterized
by relatively high energy costs due to the dissipation of thermal
energy. On the contrary, RO uses expensive osmotic membranes, which
need to be replaced regularly because of scaling and fouling. Moreover,
RO requires very high operating pressures, ranging from 20 to 200
bar, to let the water pass through the membrane. But in terms of energy
losses, it works close to the thermodynamic limit.
Thus, theoretically,
per one ion pair transferred from freshwater to salty water,
it consumes only the energy equal to the difference between the chemical potentials of the transfered ions.

The absence of large hydraulic pressure in FO process (unlike in
RO) allows to reduce the energy consumption in pumping, reduce membrane
scaling and fouling, and therefore significantly increase the lifetime
of the membranes. In FO, the draw solutes (agents) are dispersed and/or
dissolved in water form homogeneous draw solutions. The correct choice
of the draw agents is a task of paramount importance. As an osmotically
driven process, the draw solute is expected to significantly reduce
the water chemical potential, and consequently generate a high osmotic
pressure. On the other hand, the diluted draw solution is also expected
to be easily separated from fresh water and concentrated draw solution
for reuse \cite{Cai2016}.
